SoK: The Evolution of Defenses Against Web Application Attacks

### 2014-10-23 docs/questioning.tex
reshaping section 2

\title{SoK: The Evolution of Web Application Attack Defenses}

@@

In this paper we first systematize the existing knowledge
regarding the various defense mechanisms by setting up
a model for application attacks that are based on
source code injection vulnerabilities. This model
highlights the key weakness behind such attacks
and helps us observe defenses from a common
perspective.

@@

Such attacks include Cross-Site Scripting ({\sc xss})
attacks~\cite{SG07}, {\sc sql} injection attacks~\cite{RL12b},
Cross-Site Request Forgery ({\sc csrf}) attacks~\cite{LZRL09}
and others. Attacks like these top the vulnerability
lists of numerous bulletin providers like
the Open Web Application Security
Project\footnote{\url{https://www.owasp.org/index.php/Category:OWASP_Top_Ten_Project}}
({\sc owasp}) and {\sc mitre}\footnote{\url{http://cwe.mitre.org/top25/}}
for several years. Consider the case of {\sc owasp}'s Top
Ten project whose main goal is to raise awareness about
web application security by identifying some of the most
critical risks facing organizations which is referenced by
numerous researchers. In its three consecutive Top Ten lists
(2007, 2010, 2013), different source code-driven injection attacks
dominate the top five positions. This indicates that apart from
the fact that malicious users find new ways to bypass defense
mechanisms by using a variety of techniques despite the numerous
countermeasures that are being introduced.
Note that the number of systems developed to counter {\sc sql}
injection attacks until 2006 was more than twenty~\cite{HVO06}.
Since then the number has doubled.
In our research we focus on the top {\sc tba} systems that counter
application attacks, in terms of citations.
% TODO@dimitro: explain why you leave out {\it SecuriFly}~\cite{MLL05}.

@@

\section{Background}
\label{sec:attacks}

@@

\begin{figure}
\begin{center}
\leavevmode
\includegraphics[scale=0.37]{attack-tree-uml.pdf}
\end{center}
\caption{\label{fig:taxonomy}A taxonomy of code injection attacks.
In our research we have excluded the defenses that deal with
the subcategories that are illustrated in grey color.}
\end{figure}

Figure \ref{fig:taxonomy} presents a taxonomy of code injection attacks divided
into two categories. The first involves binary code and the second
executable source code.

\subsection{Binary Code Injection}

Binary code injection attacks involve
the insertion of binary code into an application
to alter its execution flow and execute malicious
compiled code. This category involves buffer-overflow
attacks~\cite{CPMHWBBGWZ98,K11}, a staple of security problems.
An extensive survey on binary code
injection attacks can be found in reference~\cite{LC03}.
Furthermore, specific advances in exploiting such vulnerabilities
(i.e. ``heap smashing", ``arc injection" and others)
have been presented in reference~\cite{PB04}.
Finally, the countermeasures used to detect
such defects have already been surveyed~\cite{YJP12}
(many of them are also included in a book:~\cite{DKZ12}
-- Section 13.8), thus we do not include them in our work.

@@

Code injection also includes the use of source code, either of a
Domain Specific Language ({\sc dsl})
or a Dynamic Language. Note that binary code injection
attacks can only occur when the target system is implemented in
languages lacking array bounds checking, like
{\sc c} and {\sc c}++. Contrariwise, source code-driven
injection attacks can target applications written in various
programming languages with different features and characteristics.

Application attacks that involve {\sc dsl}s,
constitute an important subset of the code injection problem,
as {\sc dsl}s like {\sc sql} and {\sc xml} play an
significant role in the development of either web or mobile
applications. For example, many applications have interfaces
where a user enters input to interact with the
application's data, thereby interacting with the
underlying {\it Relational Database Management System}
({\sc rdbms}).
This input can become part of an
{\sc sql} query and executed on the target {\sc rdbms}.
A code injection attack that exploits the vulnerabilities of these
interfaces by taking advantage of input validation issues like
incorrectly passed parameters or incorrect type handling,
is called an ``{\sc sql} injection
attack''~\cite{CERT02,MS09,HVO06,SW06}.
By using similar techniques malicious users
can perform other exploits based on {\sc dsl}s,
like {\sc xp}ath~\cite{SW06,CDL07,MKS09} and
{\sc xml}~\cite{MSM13}.

A recent class of {\sc cia}s involve
dynamic languages like Python, Perl, JavaScript, and {\sc php}~\cite{SFVM09,EWKK09,SMS13}.
In particular, JavaScript injection attacks comprise a wide subset of dynamic
language-driven attacks. Such attacks are manifested when a web application
accepts and redisplays data of uncertain origin without appropriate validation
and filtering. Based on this flaw, an attacker can
manage to inject a script in the JavaScript engine of a browser and alter its
execution flow~\cite{ELX07}. JavaScript injection attacks are considered as a
critical issue in web application security mainly because they are
associated with major vulnerabilities such as:
{\sc xss} attacks~\cite{SG07},
{\sc csrf} attacks~\cite{LZRL09}
and ``Cross-Channel Scripting"
({\sc xcs}) attacks~\cite{W10,BBB09}.

@@

@@

\begin{figure*}
\begin{center}
\leavevmode
\includegraphics[scale=0.53]{attacks-steps.pdf}
\end{center}
\caption{\label{fig:attacks}Attack model demonstrating web
application attacks based on code injection vulnerabilities
({\sc draft}).}
\end{figure*}

>>>>

\title{SoK: The Evolution of Defenses Against Web Application Attacks}

@@

In this paper we first systematize the existing knowledge
regarding the various defense mechanisms by setting up
a model for application attacks that are based on
executable source code injection vulnerabilities. This model
highlights the key weakness behind such attacks
and helps us observe defenses from a common
perspective.

@@

Such attacks include Cross-Site Scripting ({\sc xss})
attacks~\cite{SG07}, {\sc sql} injection attacks~\cite{RL12b},
Cross-Site Request Forgery ({\sc csrf}) attacks~\cite{LZRL09}
and others. Attacks like these top the vulnerability
lists of numerous bulletin providers like
the Open Web Application Security
Project\footnote{\url{https://www.owasp.org/index.php/Category:OWASP_Top_Ten_Project}}
({\sc owasp}) and {\sc mitre}\footnote{\url{http://cwe.mitre.org/top25/}}
for several years. Consider the case of {\sc owasp}'s Top
Ten project whose main goal is to raise awareness about
web application security by identifying some of the most
critical risks facing organizations which is referenced by
numerous researchers. In its three consecutive Top Ten lists
(2007, 2010, 2013), different executable source code-driven
injection attacks dominate the top five positions.
This indicates that apart from the fact that malicious
users find new ways to bypass defense
mechanisms by using a variety of techniques despite the numerous
countermeasures that are being introduced.
Note that the number of systems developed to counter {\sc sql}
injection attacks until 2006 was more than twenty~\cite{HVO06}.
Since then the number has doubled.
In our research we focus on the top {\sc tba} systems that counter
application attacks, in terms of citations.
% TODO@dimitro: explain why you leave out {\it SecuriFly}~\cite{MLL05}.

@@

\section{Attacks}
\label{sec:attacks}

@@

\begin{figure}
\begin{center}
\leavevmode
\includegraphics[scale=0.47]{attack-tree-uml.pdf}
\end{center}
\caption{\label{fig:taxonomy}A taxonomy of executable source
code injection attacks.}
\end{figure}

Code injection attacks can be divided into two categories.
The first involves binary code and the second executable source code.
An extensive survey on binary code
injection attacks can be found in reference~\cite{LC03}.
Furthermore, specific advances in exploiting such vulnerabilities
(i.e. ``heap smashing", ``arc injection" and others)
have been presented in reference~\cite{PB04}.
Finally, the countermeasures used to detect
such defects have already been surveyed~\cite{YJP12}
(many of them are also included in a book:~\cite{DKZ12}
-- Section 13.8), thus we do not include them in our work.
In this paper, we focus on the defenses that protect
applications against application attacks based on
the executable source code injection category.

@@

Figure~\ref{fig:taxonomy} presents a taxonomy of executable source
code injection attacks. Such attacks may involve source code, either of a
Domain Specific Language ({\sc dsl})
or a Dynamic Language. Note that binary code injection
attacks can only occur when the target system is implemented in
languages lacking array bounds checking, like
{\sc c} and {\sc c}++. Contrariwise, source code-driven
injection attacks can target applications written in various
programming languages with different features and characteristics.

Application attacks that involve {\sc dsl}s,
constitute an important subset of the code injection problem,
as {\sc dsl}s like {\sc sql} and {\sc xml} play an
significant role in the development of either web or mobile
applications. For example, many applications have interfaces
where a user enters input to interact with the
application's data, thereby interacting with the
underlying {\it Relational Database Management System}
({\sc rdbms}).
This input can become part of an
{\sc sql} query and executed on the target {\sc rdbms}.
A code injection attack that exploits the vulnerabilities of these
interfaces by taking advantage of input validation issues like
incorrectly passed parameters or incorrect type handling,
is called an {\sc sql} injection attack~\cite{CERT02,MS09,HVO06,SW06}.
By using similar techniques malicious users
can perform other exploits based on {\sc dsl}s,
like {\sc xp}ath~\cite{SW06,CDL07,MKS09} and {\sc xml}~\cite{MSM13}.
With this kind of attacks, a malicious user can view sensitive
information, destroy or modify protected data, or even crash the entire
application.

A recent class of {\sc cia}s involve
dynamic languages like Python, Perl, JavaScript, and {\sc php}~\cite{SFVM09,EWKK09,SMS13}.
JavaScript injection attacks in particular, comprise a wide subset of dynamic
language-driven attacks. Such attacks are manifested when a web application
accepts and redisplays data of uncertain origin without appropriate validation
and filtering. Based on this flaw, an attacker can
manage to inject a script in the JavaScript engine of a browser and alter its
execution flow~\cite{ELX07}. JavaScript injection attacks are considered as a
critical issue in web application security mainly because they are
associated with major vulnerabilities such as:
{\sc xss} attacks~\cite{SG07},
{\sc csrf} attacks~\cite{LZRL09}
and Cross-Channel Scripting
({\sc xcs}) attacks~\cite{W10,BBB09}.

@@

\subsection{Exploitation Model}

In this section we provide a step-by-step exploitation
model

\begin{figure*}
\begin{center}
\leavevmode
\includegraphics[scale=0.55]{attacks-steps.pdf}
\end{center}
\caption{\label{fig:attacks}Attack model demonstrating web
application attacks based on executable source
code injection vulnerabilities ({\sc draft}).}
\end{figure*}

@@
